Contextual Combinatorial Cascading Bandits

### 2015-11-30 contextual combinatorial cascading bandits.tex
add problem formulation for 'or' case

At time $t$, the learning agent is revealed with feature vectors $x_{t,a} \in \RR^{d \times 1}$ for every basic arm $a \in E$, where we have $\norm{x_{t,a}}_2 \leq 1$; this feature vector combines both information of the user and the corresponding basic arm. We use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$, which is called {\it weight} for arm $a$ at time $t$, to indicate the output of the item $a$ at time $t$. Assume $\bw_{t,a}$ are mutually independent and satisfy
\begin{equation}
\label{eq:expectation}
\EE[\bw_{t,a} | x_{t,a}] = \theta_*^{\top} x_{t,a}
\end{equation}
where $\theta_*$ is an unknown $d$-dimensional vector with the assumption that $\norm{\theta_*}_2 \leq 1$ and $0 \leq \theta_*^{\top} x_{t,a} \leq 1$ for all $t, a$. At time $t$, the learning agent chooses a solution $\bA_t=(\ba_{1}^t,...,\ba_{\abs{\bA_t}}^t) \in \cS$ based on its past observations. At the end of every time $t$, the agent observes weights of the first $\bO_t$ items of $\bA_t$ where 
$$
\bO_t \leq \abs{\bA_t}, \bw_t(\ba_k^t) = 1, \forall k < \bO_t, \bw_t(\ba_{\bO_t}^t) = 0
$$ 
or 
$$
\bO_t = \abs{\bA_t}, \bw_t(\ba_k^t) = 1, \forall k \leq \abs{\bA_t}.
$$
We say that item $a$ is {\it observed} if $a = \ba_k^t$ for some $k \leq \bO_t$.

The agent receives reward $1$ if all items in $\bA_t$ are observed and have weights $1$. Suppose we consider partial reward, $0 \leq \gamma_1 \leq \ldots \leq \gamma_{K-1} \leq 1$, for partial number of observed items with weight $1$: If $\bO_t = 1 < \abs{\bA_t}$, then the agent only observes the weight of the first item of $\bw_t(\ba_1^t)$ and receives zero reward; if $1 < \bO_t < \abs{\bA_t}$, then the agent observes weights of first $\bO_t$ items among which $\bO_t - 1$ items have weights 1 and receives reward $\gamma_{\bO_t - 1}$; if $\bO_t = \abs{\bA_t}$ and $\bw_{\ba_{\abs{\bA_t}}^t} = 0$, then the agent observes $\abs{\bA_t} - 1$ items with weights $1$ and receives reward $\gamma_{\abs{\bA_t} - 1}$. Note that the order of $\bA_t$ affects both the feedback and the reward. In the analysis part, we will show how the order affects the reward.

At the end of time $t$, the agent observes $\bO_t, \bw_t(\ba_k^t), 1\leq k\leq\bO_t$ and receives the reward
\begin{align*}
\br_t &= \begin{cases}
\gamma_{\bO_t - 1}  &\text{if } \bw_t(\ba_{\bO_t}^t) = 0\\
1 &\text{if } \bw_t(\ba_{\bO_t}^t) = 1,
\end{cases}\\
&=\begin{cases}
\gamma_{k - 1}  &\text{if } \bw_t(\ba_{i}^t) = 1, i\leq k-1, \bw_t(\ba_{k}^t) = 0, k\leq \abs{\bA_t}\\
1 &\text{if } \bw_t(\ba_{i}^t) = 1, i\leq \abs{\bA_t},
\end{cases}
\end{align*}
where we use the notation of $\gamma_0 = 0$.

Given partial reward parameters $0 \leq \gamma_1 \leq \ldots \leq \gamma_{K-1} \leq 1$, we introduce a function $f$ on $A=(a_1,...,a_{\abs{A}}) \in \cS, w=(w(1),...,w(L))$ by
\begin{align*}
&f : \cS \times [0,1]^E \to [0,1]\\
&f(A,w) = \sum_{k=2}^{\abs{A}}\gamma_{k-1} (\prod_{i=1}^{k-1}w(a_i))(1 - w(a_k)) + \prod_{i=1}^{\abs{A}}w(a_i),
\end{align*}
then we have $\br_t = f(\bA_t, \bw_t)$ and $\EE[\br_t]=f(\bA_t,\theta_*^{\top}x_t)$ where $x_t=(x_{t,1}, \cdots, x_{t,L}) \in \RR^{d \times L}$. Let 
\begin{align*}
A_t^* &= \argmax_{A\in \cS} f(A,\theta_*^{\top}x_t),\\
f_t^* &= f(A_t^*, \theta_*^{\top}x_t).
\end{align*}
The goal of our learning algorithm is to minimize the expected cumulative regret in $T$ steps
$$
R(T) = \EE[\sum_{t=1}^T \Delta_{t,\bA_t}].
$$
where $\Delta_{t,\bA_t} = f_t^* - f(\bA_t, \theta_*^{\top}x_t)$.

>>>>

At time $t$, feature vectors $x_{t,a} \in \RR^{d \times 1}$ with $\norm{x_{t,a}}_2 \leq 1$ for every basic arm $a \in E$ are revealed to the learning agent; the feature vectors combine both the information of the user and the corresponding basic arm. Then the learning agent recommends a feasible solution $\bA_t=(\ba_{1}^t,...,\ba_{\abs{\bA_t}}^t) \in \cS$ to the user. Starting from the first item, the user checks the items one by one. The checking process stops if the user clicks on one item or has checked all items without clicking anyone. We use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$ to indicate whether the user has clicked on item $a$ or not at time $t$.

Assume $\bw_{t,a}$ are mutually independent and satisfy
\begin{equation}
\label{eq:expectation}
\EE[\bw_{t,a} | x_{t,a}] = \theta_{\ast}^{\top} x_{t,a}
\end{equation}
where $\theta_{\ast}$ is an unknown $d$-dimensional vector with the assumption that $\norm{\theta_{\ast}}_2 \leq 1$ and $0 \leq \theta_{\ast}^{\top} x_{t,a} \leq 1$ for all $t, a$. At the end of every time $t$, the agent observes the first $\bO_t$ items of $\bA_t$ where 
$$
\bO_t \leq \abs{\bA_t}, \bw_t(\ba_k^t) = \begin{cases} 0, ~~k < \bO_t \\ 1, ~~k = \bO_t \end{cases}
$$ 
or 
$$
\bO_t = \abs{\bA_t}, \bw_t(\ba_k^t) = 0, ~~ k \leq \abs{\bA_t}.
$$
We say that item $a$ is {\it observed} if $a = \ba_k^t$ for some $k \leq \bO_t$.

The agent receives some reward if the user clicks on some item. Suppose we consider the position discount: if the user clicks on the $k$-th item, then the learning agent receives reward $0 \leq \gamma_k \leq 1$. Usually the importance of the positions is decreasing: the first position is most important. So it is a reasonable assumption that
$$
1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_K \geq 0.
$$
At the end of time $t$, the learning agent observes $\bO_t, \bw_t(\ba_k^t), k \leq \bO_t$ and receives reward
$$
\br_t = \bw_t(\ba_{\bO_t}^t) \gamma_{\bO_t} = \bigvee_{k=1}^{\abs{\bA_t}} \gamma_k \bw_t(\ba_k^t).
$$
where we use the notation that $\bigvee_{k=1}^n a_k = \max_{1 \leq k \leq n} a_k$. Notice that the order of $\bA_t$ affects both the feed back and the reward.

Now let us introduce a function $f$ on $A = (a_1,...,a_{\abs{A}}) \in \cS, w = (w(1),...,w(L))$ by
\begin{align*}
&f : \cS \times [0,1]^E \to [0,1]\\
&f(A,w) = \sum_{k = 1}^{\abs{A}} \gamma_{k} \prod_{i=1}^{k-1} (1 - w(a_i)) w(a_k),
\end{align*}
then we have $\br_t = f(\bA_t, \bw_t)$ and $\EE[\br_t] = f(\bA_t, \theta_{\ast}^{\top}x_t)$ where $x_t = (x_{t,1}, \cdots, x_{t,L}) \in \RR^{d \times L}$. Let 
\begin{align*}
A_t^{\ast} &= \argmax_{A\in \cS} f(A,\theta_{\ast}^{\top}x_t),\\
f_t^{\ast} &= f(A_t^{\ast}, \theta_{\ast}^{\top}x_t).
\end{align*}
The goal of our learning algorithm is to minimize the expected cumulative regret in $T$ steps
$$
R(T) = \EE[\sum_{t=1}^T R(t, \bA_t)].
$$
where $R(t, \bA_t) = f_t^{\ast} - f(\bA_t, \theta_{\ast}^{\top}x_t)$.

Similarly, we could also consider the case of conjunctive objective. We use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$ to indicate the {\it weight} of item $a$ at time $t$ satisfying equation (\ref{eq:expectation})). The learning agent observes all items until the first one with weight $0$. Suppose we also consider partial reward: if the $k$-th item is the first item with weight $0$, then the learning agent receives reward $\gamma_k'$; if all items have weight $1$, the learning agent receives reward $1$. The more items, the more reward the learning agent should receive. It is reasonable to assume that
$$
0 = \gamma_1 \leq \gamma_2 \leq \cdots \leq \gamma_K \leq 1.
$$
At the end of time $t$, the learning agent observes $\bO_t, \bw_t(\ba_k^t), k \leq \bO_t$ and receives reward
\begin{align*}
\br_{\wedge, t} &= \begin{cases}
\gamma_{\bO_t - 1}'  &\text{if } \bw_t(\ba_{\bO_t}^t) = 0,\\
1 &\text{if } \bw_t(\ba_{\bO_t}^t) = 1,
\end{cases}\\
&=\begin{cases}
\gamma_{k - 1}'  &\text{if } \bw_t(\ba_{i}^t) = 1, i < k, \bw_t(\ba_{k}^t) = 0,\\
1 &\text{if } \bw_t(\ba_{i}^t) = 1, i\leq \abs{\bA_t}.
\end{cases}
\end{align*}

If we define a function $f_{\wedge}$ on $A = (a_1, \ldots, a_{\abs{A}}) \in \cS, w = (w(1), \ldots, w(L))$ by
\begin{align*}
&f_{\wedge} : \cS \times [0,1]^E \to [0,1]\\
&f_{\wedge}(A,w) = \sum_{k = 1}^{\abs{A}} \gamma_k' (\prod_{i = 1}^{k - 1} w(a_i))(1 - w(a_k)) + \prod_{i=1}^{\abs{A}}w(a_i),
\end{align*}
then we have $\br_{\wedge, t} = f_{\wedge}(\bA_t, \bw_t)$ and $\EE[\br_{\wedge, t}] = f_{\wedge}(\bA_t, \theta_{\ast}^{\top}x_t)$ where $x_t = (x_{t,1}, \cdots, x_{t,L}) \in \RR^{d \times L}$. To simplify notations, we assume 
$$
\gamma_k' = 1 - \gamma_k.
$$
Then it holds that
\begin{equation}
\label{eq:ConDisRelation}
f_{\wedge}(A, w) = 1 - f(A, 1 - w).
\end{equation}
Let 
\begin{align*}
A_{\wedge, t}^{\ast} &= \argmax_{A\in \cS} f(A,\theta_{\ast}^{\top}x_t),\\
f_{\wedge, t}^{\ast} &= f(A_t^{\ast}, \theta_{\ast}^{\top}x_t).
\end{align*}
For the conjunctive objective, the goal of our learning algorithm is to minimize the expected cumulative regret in $T$ steps
$$
R_{\wedge}(T) = \EE[\sum_{t=1}^T R_{\wedge}(t, \bA_t)].
$$
where $R_{\wedge}(t, \bA_t) = f_{\wedge, t}^{\ast} - f_{\wedge}(\bA_t, \theta_{\ast}^{\top}x_t)$.
